Elemental Analysis of September 2013 Murrumbidgee River Water at 6 sites & 4 Tribs
==================================================================================

```{r "Set Global Options", echo=FALSE, warning=FALSE,message=FALSE}
##Firstly Set a few options
require(knitr)
options(width=200)
opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)

##Load libraries that will be used.
require(knitr)
require(psych)
require(ggplot2)
require(corrgram)
require(plyr)

## Read in base Data file
EcoChemAnalysis <- read.csv("C:/Users/s428825/Google Drive/PhD Analyses/water/data/EcoChemAnalysis.txt")

```
Have a look at the raw data:

```{r "Have a Look"}
summary(EcoChemAnalysis)
str(EcoChemAnalysis)
##describeBy(EcoChemAnalysis,EcoChemAnalysis$Location)
```

Then some data munging will be required. In this case some variables need to be designated as factors, some blank variables can be deleted entirely and NAs will be removed from variables. The code to do that is as follows:

```
##Make Location a factor Variable
EcoChemAnalysis$Location<-as.factor(EcoChemAnalysis$Location)
#Remove A couple of blank/useless variables
ECA<-within(EcoChemAnalysis, rm("AdultRecordNumber","LarvalRecordNumber"))
#Remove NAs. 
ECA<-na.omit(ECA)

##Create row names for ECA (for clustering)
row.names(ECA)<-ECA$SampleFieldLabel
#Remove junk from Field labels so thet are actual locations
ECA$SampleFieldLabel<-gsub("#.*?","",ECA$SampleFieldLabel)
ECA$Location<-gsub("#","",ECA$Location)
ECA$SampleFieldLabel<-ECA$SampleFieldLabel
ECA$SampleFieldLabel<-gsub("1.*?","",ECA$SampleFieldLabel)
ECA$SampleFieldLabel<-gsub("2.*?","",ECA$SampleFieldLabel)
#Make Factor
ECA$SampleFieldLabel<-as.factor(ECA$SampleFieldLabel)
##Set Locations as a name rather than a number.
ECA$Location<-ECA$SampleFieldLabel
##Fix some location names
ECA$Location<-gsub("umbidgee","",ECA$Location)
ECA$Location<-gsub("Xing ","Xing",ECA$Location)
ECA$Location<-gsub("Lower Molongolo ","Molongolo",ECA$Location)
```

```{r "Data Munging"}
##Make Location a factor Variable
EcoChemAnalysis$Location<-as.factor(EcoChemAnalysis$Location)
#Remove A couple of blank/useless variables
ECA<-within(EcoChemAnalysis, rm("AdultRecordNumber","LarvalRecordNumber"))
#Remove NAs. 
ECA<-na.omit(ECA)

##Create row names for ECA (for clustering)
row.names(ECA)<-ECA$SampleFieldLabel
#Remove junk from Field labels so thet are actual locations
ECA$SampleFieldLabel<-gsub("#.*?","",ECA$SampleFieldLabel)
ECA$Location<-gsub("#","",ECA$Location)
ECA$SampleFieldLabel<-ECA$SampleFieldLabel
ECA$SampleFieldLabel<-gsub("1.*?","",ECA$SampleFieldLabel)
ECA$SampleFieldLabel<-gsub("2.*?","",ECA$SampleFieldLabel)
#Make Factor
ECA$SampleFieldLabel<-as.factor(ECA$SampleFieldLabel)
##Set Locations as a name rather than a number.
ECA$Location<-ECA$SampleFieldLabel
##Fix some location names
ECA$Location<-gsub("umbidgee","",ECA$Location)
ECA$Location<-gsub("Xing ","Xing",ECA$Location)
ECA$Location<-gsub("Lower Molongolo ","Molongolo",ECA$Location)
##describeBy(ECA,ECA$Location)

##Get the locations into a sensible order. Murrumbidgee sites upstram to downstream, then the tributaries from upstream to downstream. This may be better considered as a function of distance from AC.This will require extracting the data afresh. (Al: consider use of database connect and select statement: See http://blog.yhathq.com/posts/10-R-packages-I-wish-I-knew-about-earlier.html item #5 for details)
```



```{r "Reorder by Locations"}
##Murrumbidgee on the left, tributaries on the right. Both in downstream descending order.
ECA$Location <- factor(ECA$Location, levels=c( "Angle Xing","Tharwa","Murr@TuggC","Murr@Cotter","Uriarra","Taemas","Gudgenby","Tugg.Crk","Cotter","Molongolo"), ordered=TRUE)

```
Sample Histograms
-----------------
Do some basic plots to get a feel for the data. Start with histograms. Are they normally distributed?

```{r "Basic Histograms"}
hist(ECA$Strontium, main="Histogram of Strontium")
hist(ECA$Barium)
hist(ECA$Magnesium)
hist(ECA$Manganese)
##Or could use ggplot
##MnHist <- ggplot(ECA, aes(x=Manganese))
##MnHist + geom_histogram(binwidth = 3)+labs(title = "Manganese")+ labs(y = "Ratio")+ theme(axis.text.x=element_text(angle=90))
hist(ECA$Boron)
hist(ECA$Iron)
hist(ECA$Silicon)
hist(ECA$Aluminium)
hist(ECA$Sulphur)
hist(ECA$Chlorine)
```

Elements as they Vary Between Sites
-----------------------------------
Elements as they vary with Location. The Murrumbidgee sites are the 6 on the left and in descending order. The remaining 4 on the right hand side of the graphs are the 4 tributaries. Gudgenby River, Tuggeranong Creek, Cotter River and the Molongolo River. These are also in descending order.    

```{r "Plots of Main Five"}
plot(ECA$Location, ECA$Strontium, main="Strontium",las=2, ylab="Concentration (ug/g)")
plot(ECA$Location, ECA$Barium, main="Barium", las=2, ylab="Concentration (ug/g)")
plot(ECA$Location, ECA$Magnesium, main="Magnesium",las=2, ylab="Concentration (ug/g)")
plot(ECA$Location, ECA$Manganese, main="Manganese", las=2, ylab="Concentration (ug/g)")
plot(ECA$Location, ECA$Boron, main="Boron", las=2, ylab="Concentration (ug/g)")
plot(ECA$Location, ECA$Iron, main="Iron", las=2, ylab="Concentration (ug/g)")
plot(ECA$Location, ECA$Silicon, main="Silicon", las=2, ylab="Concentration (ug/g)")
plot(ECA$Location, ECA$Aluminium, main="Aluminium", las=2, ylab="Concentration (ug/g)")
plot(ECA$Location, ECA$Sulphur, main="Sulphur",las=2, ylab="Concentration (ug/g)")
plot(ECA$Location, ECA$Chlorine, main="Chlorine", las=2, ylab="Concentration (ug/g)")
```
Ratios of the Big Four with Calcium as they Vary Between Sites
--------------------------------------------------------------

Basic plots of main elements of interest as ratios with Calcium.

```{r "Some Ratio Plots"}
SrCa <- ggplot(ECA, aes(Location,Sr.Ca))
SrCa + geom_point(alpha = 2/3) +geom_boxplot()+labs(title = "Strontium:Calcium Ratio")+ labs(y = "Ratio")+ theme(axis.text.x=element_text(angle=90))

BaCa <- ggplot(ECA, aes(Location,Ba.Ca))
BaCa + geom_point(alpha = 2/3) +geom_boxplot()+labs(title = "Barium:Calcium Ratio")+ labs(y = "Ratio")+ theme(axis.text.x=element_text(angle=90))

MgCa <- ggplot(ECA, aes(Location,Mg.Ca))
MgCa + geom_point(alpha = 2/3) +geom_boxplot()+labs(title = "Magnesium:Calcium Ratio")+ labs(y = "Ratio")+ theme(axis.text.x=element_text(angle=90))

MnCa <- ggplot(ECA, aes(Location,Mn.Ca))
MnCa + geom_point(alpha = 2/3) +geom_boxplot()+labs(title = "Manganese:Calcium Ratio")+ labs(y = "Ratio")+ theme(axis.text.x=element_text(angle=90))
```

Correlation Matrix
------------------
Plot a simple correlation Matrix to see if there is anything else of interest.

```{r "Correlation Matrix)"}
ECAcorrgram<-within(ECA, rm("ID"))
corrgram(x=ECAcorrgram)
```


Clusters - Levels and then Ratios
---------------------------------
Are clusters evident based on the levels of the four elements (sr,Ba, Mg,Mn) in the water at each site even during this period of moderate flow?

```{r "clustering of various measures" }

#See if there is clustering around raw levels
df<-subset(ECA, select = c(Strontium,Barium,Magnesium,Manganese))
hc<-hclust(dist(df))
plot(hc, main="Strontium,Barium,Calcium,Manganese", hang = -1)


##What if we include calcium in this?
df<-subset(ECA, select = c(Strontium,Barium,Magnesium,Manganese,Calcium))
hc<-hclust(dist(df))
plot(hc,main= "Including Calcium with Strontium,Barium,Calcium,Manganese", hang = -1)
```

What if we include more elements?
---------------------------------
```{r "More Elements"}
df<-subset(ECA, select = c(Strontium,Barium,Magnesium,Manganese,Calcium, Boron, Chlorine, Sulphur, Silicon, Aluminium,Iron))
hc<-hclust(dist(df))
plot(hc,main= "Including Previous 5 and Boron, Chlorine, Sulphur, Silicon, Aluminium, Iron", hang = -1)
```


Are clusters evident based on the ratio of these four elements with Calcium?

```{r "Clustering of Ratios"}
#See if there is clustering around ratios
dfr<-subset(ECA, select = c(Sr.Ca,Ba.Ca,Mn.Ca,Mg.Ca))
hcr<-hclust(dist(dfr))
plot(hcr, main="Ratios of same 4 elements with Ca", hang = -1)
```


Table of Original Data
----------------------
```{r "Print Original Data to Share"}
##Subset for Sharing
ECAshare<-ECA[,7:26]
ECAshare<-ECAshare[,-3]
ECAshare<-ECAshare[,-1]
##Relabel some remaining variables suitable for sharing
ECAshare <- rename(ECAshare, replace=c("EcoChemLabNumber" = "LabNo"))

print(ECAshare)
```
The 'r' code chunks that are in this document are:
```{r "Include Labels for all code Chunks"}
all_labels()
```