Elemental Analysis of September 2013 Murrumbidgee River Water at 6 sites & 4 Tribs
==================================================================================

```{r "Set Global Options", echo=FALSE, warning=FALSE,message=FALSE}

#The following 2 and 4 lines are needed if knitr is to work with ProjectTemplate.
require(knitr)
if (basename(getwd()) == "src") setwd("..") #needed to get knitr to work with project template
library('ProjectTemplate') #All projectTemplates need this up front
load.project() #All projectTemplates need this up front

##Firstly Set a few options
options(width=200)
opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```
Sample Histograms
-----------------
Do some basic plots to get a feel for the data. Start with histograms. Are they normally distributed?

```{r "Basic Histograms"}
hist(ECA$Strontium, main="Histogram of Strontium")
hist(ECA$Barium)
hist(ECA$Magnesium)
hist(ECA$Manganese)
##Or could use ggplot
##MnHist <- ggplot(ECA, aes(x=Manganese))
##MnHist + geom_histogram(binwidth = 3)+labs(title = "Manganese")+ labs(y = "Ratio")+ theme(axis.text.x=element_text(angle=90))
hist(ECA$Boron)
hist(ECA$Iron)
hist(ECA$Silicon)
hist(ECA$Aluminium)
hist(ECA$Sulphur)
hist(ECA$Chlorine)
```

Elements as they Vary Between Sites
-----------------------------------
Elements as they vary with Location. The Murrumbidgee sites are the 6 on the left and in descending order. The remaining 4 on the right hand side of the graphs are the 4 tributaries. Gudgenby River, Tuggeranong Creek, Cotter River and the Molongolo River. These are also in descending order.    

```{r "Plots of Main Five"}
plot(ECA$Location, ECA$Strontium, main="Strontium",las=2, ylab="Concentration (ug/g)")
plot(ECA$Location, ECA$Barium, main="Barium", las=2, ylab="Concentration (ug/g)")
plot(ECA$Location, ECA$Magnesium, main="Magnesium",las=2, ylab="Concentration (ug/g)")
plot(ECA$Location, ECA$Manganese, main="Manganese", las=2, ylab="Concentration (ug/g)")
plot(ECA$Location, ECA$Boron, main="Boron", las=2, ylab="Concentration (ug/g)")
plot(ECA$Location, ECA$Iron, main="Iron", las=2, ylab="Concentration (ug/g)")
plot(ECA$Location, ECA$Silicon, main="Silicon", las=2, ylab="Concentration (ug/g)")
plot(ECA$Location, ECA$Aluminium, main="Aluminium", las=2, ylab="Concentration (ug/g)")
plot(ECA$Location, ECA$Sulphur, main="Sulphur",las=2, ylab="Concentration (ug/g)")
plot(ECA$Location, ECA$Chlorine, main="Chlorine", las=2, ylab="Concentration (ug/g)")
```
Ratios of the Big Four with Calcium as they Vary Between Sites
--------------------------------------------------------------

Basic plots of main elements of interest as ratios with Calcium.

```{r "Some Ratio Plots"}
SrCa <- ggplot(ECA, aes(Location,Sr.Ca))
SrCa + geom_point(alpha = 2/3) +geom_boxplot()+labs(title = "Strontium:Calcium Ratio")+ labs(y = "Ratio")+ theme(axis.text.x=element_text(angle=90))

BaCa <- ggplot(ECA, aes(Location,Ba.Ca))
BaCa + geom_point(alpha = 2/3) +geom_boxplot()+labs(title = "Barium:Calcium Ratio")+ labs(y = "Ratio")+ theme(axis.text.x=element_text(angle=90))

MgCa <- ggplot(ECA, aes(Location,Mg.Ca))
MgCa + geom_point(alpha = 2/3) +geom_boxplot()+labs(title = "Magnesium:Calcium Ratio")+ labs(y = "Ratio")+ theme(axis.text.x=element_text(angle=90))

MnCa <- ggplot(ECA, aes(Location,Mn.Ca))
MnCa + geom_point(alpha = 2/3) +geom_boxplot()+labs(title = "Manganese:Calcium Ratio")+ labs(y = "Ratio")+ theme(axis.text.x=element_text(angle=90))
```

Correlation Matrix
------------------
Plot a simple correlation Matrix to see if there is anything else of interest.

```{r "Correlation Matrix)"}
ECAcorrgram<-within(ECA, rm("ID"))
corrgram(x=ECAcorrgram)
```


Clusters - Levels and then Ratios
---------------------------------
Are clusters evident based on the levels of the four elements (sr,Ba, Mg,Mn) in the water at each site even during this period of moderate flow?

```{r "clustering of various measures" }

#See if there is clustering around raw levels
df<-subset(ECA, select = c(Strontium,Barium,Magnesium,Manganese))
hc<-hclust(dist(df))
plot(hc, main="Strontium,Barium,Calcium,Manganese", hang = -1)


##What if we include calcium in this?
df<-subset(ECA, select = c(Strontium,Barium,Magnesium,Manganese,Calcium))
hc<-hclust(dist(df))
plot(hc,main= "Including Calcium with Strontium,Barium,Calcium,Manganese", hang = -1)
```

What if we include more elements?
---------------------------------
```{r "More Elements"}
df<-subset(ECA, select = c(Strontium,Barium,Magnesium,Manganese,Calcium, Boron, Chlorine, Sulphur, Silicon, Aluminium,Iron))
hc<-hclust(dist(df))
plot(hc,main= "Including Previous 5 and Boron, Chlorine, Sulphur, Silicon, Aluminium, Iron", hang = -1)
```


Are clusters evident based on the ratio of these four elements with Calcium?
```{r "Clustering of Ratios"}
#See if there is clustering around ratios
dfr<-subset(ECA, select = c(Sr.Ca,Ba.Ca,Mn.Ca,Mg.Ca))
hcr<-hclust(dist(dfr))
plot(hcr, main="Ratios of same 4 elements with Ca", hang = -1)
```


Table of Original Data
----------------------
```{r "Print Original Data to Share"}
##Subset for Sharing
ECAshare<-ECA[,7:26]
ECAshare<-ECAshare[,-3]
ECAshare<-ECAshare[,-1]
##Relabel some remaining variables suitable for sharing
ECAshare <- rename(ECAshare, replace=c("EcoChemLabNumber" = "LabNo"))

print(ECAshare)
```
The 'r' code chunks that are in this document are:
```{r "Include Labels for all code Chunks"}
all_labels()
```